Three magnified views of a supernova explosion 10 billion years ago, and a fourth that will reveal the nature of cosmological dark energy

### 2020-03-28 Paper/Nature/main.tex
Presubmission to Nature

\begin{abstract}
% https://www.nature.com/documents/nature-summary-paragraph.pdf
%One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline
The expansion history of the Universe is one of the most fundamental observations leading to the hot big bang model for the origin of the Universe. 
%Two to three sentences of more detailed background, comprehensible to scientists in related disciplines.
The observed brightness of distant Type Ia supernova (SN) explosions has played a key role in mapping the expansion history, and led to the discovery of dark energy that now appears to be driving an accelerating cosmic expansion rate.
Determining the nature of dark energy and how it may evolve over time is a primary goal for current and planned large-scale cosmology experiments.
%One sentence clearly stating the general problem being addressed by this particular study
One promising method for determining the dark energy equation of state and its evolution  is through time delays of gravitational lens systems, which for a given lens potential provide a measurement of the ratio of the angular diameter distances to the foreground lens and background source.
%One sentence summarising the main result (with the words “here we show” or their equivalent).
Here we report the discovery of a strongly-lensed Type Ia SN explosion that will provide a time delay precision of $<1\%$.  The SN is in an evolved galaxy at $z=1.95$ gravitationally lensed by a massive foreground galaxy cluster.
%Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
With three lensed images of the supernova detected in existing observations from the \textit{Hubble Space Telescope} with time delays of order 50 days, this is the most distant multiply-imaged supernova yet discovered.  A fourth image in the component closest to the cluster core is predicted to appear in the year 2035$\pm$2.  Observing the lightcurve of that future image will provide a time delay precision of $\approx 7$ days over an unprecedented baseline of nearly 20 years.  This single source will therefore provide a similar precision and complementary cosmological constraints to an ensemble of dozens of time delay measurements expected at somewhat lower redshifts and shorter timescales from the planned ground-based LSST and WFIRST space telescope missions.  If we suppose that cosmological parameters such as the local expansion rate ($H_0$) and dominant energy densities $\Omega_m$ and $\Omega_\Lambda$ will be well established in 2035, the combined time delay measurements will constrain the dark energy equation of state ($w_0$) to better than 10\%.
%One or two sentences to put the results into a more general context. 
% Considering this observation in the context of
% Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{../Figures/snrequiem_w0wa_compared_to_lsst_wfirst.pdf}
    \caption{The unusual sensitivity of SN Requiem to dark energy equation of state parameters ($w_0, w_a$). Contours show the 1$\sigma$ confidence region for a typical sample of lensed SNIa time delays from LSST (blue) and WFIRST (red) compared to SN Requiem (magenta). Dashed lines and the white marker show the input values of $w_0=-1$, $w_a=0$ that were used for this simulation. In order to isolate the $w_0, w_a$ dependence, all contours are constructed assuming perfect knowledge of $H_0$, $\Omega_{m,0}$ and $\Omega_{\rm de,0}$.  SN Requiem could be highly complementary to the LSST+WFIRST lensed SN sample, due to its unusual combination of a relatively low-redshift lens ($z=0.338$) and high-redshift source ($z=1.95$). This single SN could deliver constraints on dark energy equation of state parameters comparable to several dozen SNe with shorter and less precise  time delays.}
    \label{fig:my_label}
\end{figure}

>>>>

\begin{abstract}
% https://www.nature.com/documents/nature-summary-paragraph.pdf
%One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline
The expansion history of the Universe is one of the most fundamental observations leading to the hot big bang model for the origin of the Universe\footnote{Penzias \& Wilson 1965, ApJL, 142, 419}. 
%Two to three sentences of more detailed background, comprehensible to scientists in related disciplines.
The observed brightness of distant Type Ia supernova (SN) explosions has played a key role in mapping the expansion history, and led to the discovery of dark energy that now appears to be driving an accelerating cosmic expansion rate\footnote{Riess et al. 1998, AJ, 116, 1009; Perlmutter et al. 1999, ApJ, 517, 565; Riess et al. 2007, ApJ, 659, 98}.
Determining the nature of dark energy and how it may evolve over time is a primary goal for current and planned large-scale cosmology experiments\footnote{Amendola et al. 2013, Living Reviews in Relativity, Volume 16, Issue 1, article id. 6, 270 pp.; Ivezic et al. 2019, ApJ, 873, 111}.
%One sentence clearly stating the general problem being addressed by this particular study
One promising method for determining the dark energy equation of state and its evolution  is through time delays of gravitational lens systems, which for a given lens potential provide a measurement of the ratio of the angular diameter distances to the foreground lens and background source\footnote{Coe \& Moustakas, 2009, Apj 706, 45; Suyu 2010, ApJ, 711, 201}.
%One sentence summarising the main result (with the words “here we show” or their equivalent).
Here we report the discovery of a strongly-lensed Type Ia SN explosion that will provide a time delay precision of $<1\%$.  The SN is in an evolved galaxy at $z=1.95$ gravitationally lensed by a massive foreground galaxy cluster.
%Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
With three lensed images of the supernova detected in existing observations from the \textit{Hubble Space Telescope} with time delays of order 50 days, this is the most distant multiply-imaged supernova yet discovered.  A fourth image in the component closest to the cluster core is predicted to appear in the year 2035$\pm$2.  Observing the lightcurve of that future image will provide a time delay precision of $\approx 7$ days over an unprecedented baseline of nearly 20 years.  This single source will therefore provide a similar precision and complementary cosmological constraints to an ensemble of dozens of time delay measurements expected at somewhat lower redshifts and shorter timescales from the planned ground-based LSST and WFIRST space telescope missions.  If we suppose that cosmological parameters such as the local expansion rate ($H_0$) and dominant energy densities $\Omega_m$ and $\Omega_\Lambda$ will be well established in 2035, the combined time delay measurements will constrain the dark energy equation of state ($w_0$) to a precision of 20\%.
%One or two sentences to put the results into a more general context. 
% Considering this observation in the context of
% Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{../Figures/snrequiem_w0wa_compared_to_lsst_wfirst.pdf}
    \caption{The unusual sensitivity of SN Requiem to dark energy equation of state parameters ($w_0, w_a$). Contours show the 1$\sigma$ confidence region for a typical sample of lensed SNIa time delays from LSST (blue) and WFIRST (red) compared to SN Requiem (magenta), which is shown for assumed uncertainties on the lens potential of 5\% and 2\%. Dashed lines and the white marker show the input values of $w_0=-1$, $w_a=0$ that were used for this simulation. In order to isolate the $w_0, w_a$ dependence, all contours are constructed assuming perfect knowledge of $H_0$, $\Omega_{m,0}$ and $\Omega_{\rm de,0}$.  SN Requiem could be highly complementary to the LSST+WFIRST lensed SN sample, due to its unusual combination of a relatively low-redshift lens ($z=0.338$) and high-redshift source ($z=1.95$). This single SN could deliver constraints on dark energy equation of state parameters comparable to several dozen SNe with shorter and less precise  time delays.}
    \label{fig:my_label}
\end{figure}
